---
title             : "Aggregation experiment summary"
shorttitle        : ""

note              : Created `r format(Sys.time())`

author: 
  - name          : "Shir Dekel"
  
floatsintext      : yes

bibliography      : '`r system.file("refs.bib", package = "shiR")`'
csl               : '`r system.file("apa.csl", package = "shiR")`'
documentclass     : "apa7"
classoption       : "man, donotrepeattitle"
output: 
  papaja::apa6_pdf:
    includes: 
      in_header: "preamble.tex"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# Background

- An aggregated portfolio of (positive EV)[^1] gambles has lower overall risk and lower probability of making an overall loss than any individual gamble.
- Choice bracketing (grouping choices) seems to facilitate EV maximisation (accept positive EV gambles and reject negative EV gambles).
  - Suggests that choice bracketing might facilitate aggregation.

[^1]: Note that the gambles referred to in this memo will be assumed to have a positive EV, unless otherwise specified.

## Evidence of choice bracketing facilitating EV maximisation

- Multi-play gambles
  - People accept a single decision about multiple identical gambles more than a single decision about one of those gambles [@wedell1994; @dekay2005; @keren1991; @montgomery1982; @redelmeier1992].
  - People accept a single decision about multiple identical gambles more with an outcome probability distribution, than without [@benartzi1999; @redelmeier1992; @klos2013; @webb2017; @coombs1971; @venkatraman2006; @dekay2005; @langer2001; @keren1991].
- Repeated-play gambles (experience-based)
  - People accept a series of gambles more when are they are evaluated and decisions are made less frequently [@gneezy1997; @thaler1997; @bellemare2005; @beshears2016].
  - People accept a series of gambles more when when they receive feedback or are able to sample from a distribution before making a choice [@camilleri2011; @camilleri2013; @barron2003; @wulff2018; @ludvig2011; @hertwig2004; @jessup2008].
  - People accept a series of gambles more when being instructed to consider the options as a part of a portfolio [@sokolhessner2009; @sokolhessner2012].
- Repeated-play gambles (description-based)
  - People choose fewer (negative EV) lottery tickets when considering them jointly, than when considering them one at a time [@haisley2008].

## Limitations

- Multi-play gambles
  - Risky choices are usually evaluated in separate instances, not all as a part of one decision.
  - Choices are usually not identical.
  - The outcome distribution is rarely available.
- Repeated-play gambles (experience-based)
  - We rarely see the immediate outcomes of risky choices, or are able to sample from the outcome distribution.
- Repeated-play gambles (description-based)
  - Outcomes and probabilities of each gamble were not specified.
  - Not a good explanation about why this occurs.

## Choice bracketing manipulations for description-based repeated-play gambles

- Similarity
  - Choices whose differences are easy to compare (alignable differences) are weighted heavier than those that are difficult to compare [@markman1995; @markman2010].
  - As such, gambles that are more similar are likely to be broadly bracketed.
- Presentation
  - Attributes that are difficult to evaluate on their own will have a greater impact on joint presentation than separate presentation [@hsee1999].
  - As such, gambles that are presented jointly are likely to be broadly bracketed.
- Awareness
  - Lengthy instructions with an elaborate explanation to "think like a trader" -- and to consider all the repeated-play gambles as a portfolio, as opposed to considering them individually -- reduced risk aversion [@sokolhessner2009; @sokolhessner2012].
  - As such, gambles are likely to be broadly bracketed when considered as a portfolio.

# Experiment

- Research question: how can we facilitate risk aggregation in repeated gambles?
- Hypotheses
  - Participants that know how many projects to expect will make more risky choices than participants that are unaware.
  - Participants will make more risky choices when seeing projects jointly than when seeing them separately.
  - Participants that see projects from the same industry will make more risky choices than participants that see projects from different industries.
- Novelty
  - More isomorphic with business: independent decisions without immediate feedback.
  - New choice bracketing manipulations: similarity and awareness.

## Method

### Participants

- N = 198
- From Prolific
- Age range: 18-69 
- Age mean: 32.52 (SD = 11.45)
- Business experience mean years: 7.01 (median = 3)
- Business education mean years: 1.70 (median = 0)

### Design

- IVs
  - Similarity (between-subjects)
  - Awareness (between-subjects)
  - Presentation (within-subjects)
- DV
  - Proportion of projects accepted

### Materials

Figure\ \@ref(fig:instructions) shows the instructions. Figures\ \@ref(fig:awareness-aware) and\ \@ref(fig:awareness-naive) show the next webpage shown to those in the aware and naive conditions, respectively. Figures\ \@ref(fig:presentation-separate-similarity-high1) and\ \@ref(fig:presentation-separate-similarity-high2) show examples of the following 10 separate gambles, in the high similarity condition. Figures\ \@ref(fig:presentation-separate-similarity-low1) and\ \@ref(fig:presentation-separate-similarity-low2) show the same, but in the low similarity condition. Figures\ \@ref(fig:presentation-joint-similarity-high) and\ \@ref(fig:presentation-joint-similarity-low) show a (cropped) set of the 10 joint gambles in the high and low similarity conditions, respectively. Figure\ \@ref(fig:distribution) shows the question with the outcome distribution shown to all participants after the repeated gambles.

```{r instructions, out.width = "100%", fig.cap = "Instructions", fig.pos = "H"}
image_read(here("img","instructions.png")) %>% 
        image_trim()
```

```{r awareness-aware, out.width = "100%", fig.cap = "Awareness - Aware condition", fig.pos = "H"}
image_read(here("img","awareness_aware.png")) %>% 
        image_trim()
```

```{r awareness-naive, out.width = "20%", fig.cap = "Awareness - Naive condition", fig.pos = "H"}
image_read(here("img","awareness_naive.png")) %>% 
        image_trim()
```

```{r presentation-separate-similarity-high1, out.width = "100%", fig.cap = "Separate presentation - High Similarity example 1", fig.pos = "H"}
image_read(here("img","presentation_separate_similarity_high1.png")) %>% 
        image_trim()
```

```{r presentation-separate-similarity-high2, out.width = "100%", fig.cap = "Separate presentation - High Similarity example 2", fig.pos = "H"}
image_read(here("img","presentation_separate_similarity_high2.png")) %>% 
        image_trim()
```

```{r presentation-separate-similarity-low1, out.width = "100%", fig.cap = "Separate presentation - Low Similarity example 1", fig.pos = "H"}
image_read(here("img","presentation_separate_similarity_low1.png")) %>% 
        image_trim()
```

```{r presentation-separate-similarity-low2, out.width = "100%", fig.cap = "Separate presentation - Low Similarity example 2", fig.pos = "H"}
image_read(here("img","presentation_separate_similarity_low2.png")) %>% 
        image_trim()
```

```{r presentation-joint-similarity-high, out.width = "100%", fig.cap = "Joint presentation - High similarity (cropped for space)", fig.pos = "H"}
image_read(here("img","presentation_joint_similarity_high.png")) %>% 
  image_trim() %>%
  image_crop("1485x1880+0+0")
```

```{r presentation-joint-similarity-low, out.width = "100%", fig.cap = "Joint presentation - Low similarity (cropped for space)", fig.pos = "H"}
image_read(here("img","presentation_joint_similarity_low.png")) %>% 
  image_trim() %>%
  image_crop("1485x1880+0+0")
```

```{r distribution, out.width = "100%", fig.cap = "Distribution", fig.pos = "H"}
image_read(here("img","distribution.png")) %>% 
        image_trim()
```

## Results

- Figure\ \@ref(fig:full) shows all three variables
- Participants invested more when they were told that there will be 10 projects, compared to when they were not told this
  - See Figure\ \@ref(fig:awareness)
- Participants invested more when viewing the projects jointly, compared to when they viewed them separately
  - See Figure\ \@ref(fig:presentation)
- There was no main effect of similarity
- The interaction between similarity and presentation was significant
  - See Figure\ \@ref(fig:similarity-presentation)
  - The presentation effect seems stronger in the high similarity condition than in the low similarity condition
- An analysis of the trial-by-trial data showed that those in the aware condition became riskier over the 10 separate trials, but those in the naive condition did not.
  - See Figure\ \@ref(fig:plot-separate-trial-awareness).
  
```{r full, fig.cap = "Mean proportions of decisions to invest in each set of 10 projects, by presentation, similarity, and awareness conditions. Error bars represent 95% confidence intervals."}
readd(full)
```

```{r awareness, fig.cap = "Mean proportions of decisions to invest in each set of 10 projects, by awareness conditions. Error bars represent 95% confidence intervals."}
readd(awareness)
```

```{r presentation, fig.cap = "Mean proportions of decisions to invest in each set of 10 projects, by presentation conditions. Error bars represent 95% confidence intervals."}
readd(presentation)
```

```{r similarity-presentation, fig.cap = "Mean proportions of decisions to invest in each set of 10 projects, by presentation and similarity conditions. Error bars represent 95% confidence intervals."}
readd(similarity_presentation)
```

```{r plot-separate-trial-awareness, fig.cap = "Proportion of project acceptance in the separate presentation condition, by trial and awareness conditions. The smoothing over trials is done using LOESS, and the shading represents 95% confidence intervals."}
readd(plot_separate_trial_awareness)
```

# Limitations

- Outcome distribution
  - The outcome distribution (presented after the gambles) was flipped such that it appeared more attractive than the actual distribution underlying the gambles.
- Similarity counterbalancing
  - I didn't counterbalance the project domains, such that in the high similarity condition all participants saw the oil well project.
  - There is a chance this will explain some of the interaction effect, since perhaps the nature of the project domain may have influenced the risk taking somehow.
- The similarity x presentation interaction
  - We did not predict this interaction and struggled to explain it.
  - Also, it isn't a massive effect, so it may be spurious.
- Presentation counterbalancing
  - There is an awareness effect in the joint presentation condition.
  - This is odd, since participants can clearly see the number of gambles.
  - Might be an order effect.

# Follow up

## Changes

Distribution

- Use a corrected distribution
- Manipulate whether or not participants see a distribution alongside their gambles.
  - Will allow to test whether people are aggregating naively or normatively.
  - If there's an effect of distribution (which we'd hypothesise), then participants are not aggregating perfectly.
  - Alongside an effect of presentation/awareness would suggest an aggregation "heuristic".

Similarity

- It is unclear what caused the similarity simple effect we found.
- There is reason to believe that it is a naive diversification effect (people preferring heterogeneity).
- However, it is unclear what would be a way to separate this from a potential bracketing effect (homogeneity facilitating bracketing).
  - Both could be contributing.
- Should probably remove this manipulation for the follow-up.

Presentation

- Manipulate between-subjects in order to rule out order effects.

Awareness
  
- We can add a question after the gambles for the naive condition to ask how many gambles they think they saw.
  - Might help understand the effect.

Business experience

- We should add more nuanced demographics question that queries participants' business experience and education.
  - At the moment it's just a question about how many years of experience/education in business they have.
  - It seems important, especially since the prolific sample reported quite a high average years of experience.
  - Asking these of the laypeople would also allow for a more direct comparison with the managers later on.
- Specifically:
  - Company revenue
  - Role at the company
  - Budget under control

Chance percentages

- According to Dan, percentage chance values that are not rounded to the nearest 5%, such as 26% like in our materials, are uncommon in the business world.

## Design

- We would be comparing three main effects in a non-factorial design:
  - Presentation: joint vs. separate (in the naive awareness condition)
  - Awareness: aware vs naive (in the separate presentation condition)
  - Distribution: distribution vs no distribution (for all conditions)
- Hypotheses (see Figure\ \@ref(fig:plot-predictions)):
  - Joint Naive will be higher than Separate Naive (presentation effect).
  - Separate Aware will be higher than Separate Naive (awareness effect).
  - Distribution will be higher than no distribution (distribution effect).

```{r plot-predictions, fig.cap = "Predictions for the aggregation follow-up experiment."}
readd(plot_predictions)
```

\newpage

# References

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
